Towards a Complete Model of Mutation Selection Balance in Humans

### 2023-05-13 manuscript/manuscript.tex
more theory section stuff

When mutations are weakly deleterious, their allele frequencies are strongly
influenced by stochastic perturbations and their frequency trajectories are no
longer well-approximated by deterministic models. When genetic drift is the
only source of randomness in allele frequency change, the relevant scale of
stochastic perturbations is determined by the drift-effective population size
$N_d$ (e.g. due to only non-selective process). For reasons that will soon be
clear, we consider a general expression of the drift-effective population size
that allows for extra non-heritable variability in offspring number $V_k$,

\begin{align}
    N_d = \frac{2N}{V_k + 2}
\end{align}

@@

To further complicate matters, both strong and weakly deleterious mutations
create \emph{heritable} fitness variation. In turn, the presence of heritable
fitness variation generates another source of randomness in allele frequency
change known as genetic \emph{draft} \parencite{Neher2013-dz}. Genetic draft
occurs when alleles become randomly associated with different fitness
backgrounds each generation, which stochastically perturbs their trajectories.
Like genetic drift, perturbations due to draft are directionless but overall
increase the variance in allele frequency change. Unlike drift, the stochastic
perturbations each generation due to draft are autocorrelated as long as
segregation and recombination do not decouple an allele from its fitness
background and the selective environment does not change
\parencite{Robertson1961-ho,Santiago1995-hx,Buffalo2019-qs}. When selection is
strong,

When selection is
weak and fitness is polygenic, the stochasticity introduced by draft is
well-approximated by rescaling the drift-effective population size

that exist when there are heritable differences among haplotypes.
Mathematical models of genetic draft

Gillespie2000-mh

,Robertson1961-ho,Santiago1995-hx,Santiago1998-bs,Barton2000-zg

\subsection*{DFE Estimates}
 - Table s5 https://oup.silverchair-cdn.com/oup/backfile/Content_public/Journal/genetics/206/1/10.1534_genetics.116.197145/8/files1.pdf?Expires=1686357613&Signature=cp9qoI2ChCb6GFZ74HI4eWcA~MQ4sHlWyeI3Zp49-5xEORzMfYlSRxzVvZlo8ZoGWDi0G3PnL00ZplM-~or4YIVm2uhkmqjf61AmC7RdZ1QTak4j0kFMQ0wAbM9pusiV3gIepTQHq2HKRKwgS~ypmv~0oqO5dDxaq9nhOWIBBmLAA49oHWuxUtnRe3sEYATwCwjIFEomhgJpg7vKGCF~87GdCV4JETW4PN2VvD4hsu2AVRd8yA~vUJvsXfhWwcElDeY4fsgpprl1lC3XGjJCnqdwbzRGPV5LNx5XEKLptcZtLeLcOmFJCvmVLe0bewJ7Vta0TSQM7T9AuklompjPPw__&Key-Pair-Id=APKAIE5G5CRDK6RD3PGA

>>>>

When mutations are weakly deleterious, their allele frequencies are strongly
influenced by stochastic perturbations and their frequency trajectories are no
longer well-approximated by deterministic models. When genetic drift is the
only source of randomness in allele frequency change, the relevant scale of
stochastic perturbations is determined by the drift-effective population size
$N_d$ (e.g. due to only non-selective process). For reasons that will soon be
clear, we consider a general expression of the drift-effective population size
that allows for extra non-heritable variability in offspring number $V_k$,
%
\begin{align}
    N_d = \frac{2N}{V_k + 2}
\end{align}

@@

To further complicate matters, both strong and weakly deleterious mutations
create \emph{heritable} fitness variation. In turn, the presence of heritable
fitness variation generates another source of randomness in allele frequency
change known as genetic \emph{draft} \parencite{Neher2013-dz}. Genetic draft
occurs when alleles become randomly associated with different fitness
backgrounds each generation, which stochastically perturbs their trajectories.
Like genetic drift, perturbations due to draft are directionless but overall
increase the variance in allele frequency change. Unlike drift, stochastic
perturbations caused by draft are autocorrelated across generations as long an
allele does not recombine or segregate off its fitness background and the
selective environment does not change
\parencite{Robertson1961-ho,Santiago1995-hx,Buffalo2019-qs}. When selection is
strong, draft can generate multiple-merger coalescences that create
discontinuities in allele frequency trajectories, making diffusion
approximations ill-suited \parencite{Gillespie2000-mh,Der2011-it,Neher2013-dz}.
However, when selection is weak and fitness is polygenic, the stochasticity
introduced by draft can be approximated by assuming draft just increases the
rate of drift. This is often expressed as rescaling the effective population
size

Though there are subtle differences in the effective population sizes

rescaling the
drift-effective population size by an appropriate factor
\parencite{Robertson1961-ho,Santiago1995-hx,Santiago1998-bs,Barton2000-zg}.
While a simple rescaling cannot elucidate details about how selection alters
the genealogy, it does allow for the calculation of key quantities like the
pairwise coalescence time (though see \textcite{Santiago1998-bs}, p. 2111).

\subsection*{DFE Estimates}
% - Table s5 https://oup.silverchair-cdn.com/oup/backfile/Content_public/Journal/genetics/206/1/10.1534_genetics.116.197145/8/files1.pdf?Expires=1686357613&Signature=cp9qoI2ChCb6GFZ74HI4eWcA~MQ4sHlWyeI3Zp49-5xEORzMfYlSRxzVvZlo8ZoGWDi0G3PnL00ZplM-~or4YIVm2uhkmqjf61AmC7RdZ1QTak4j0kFMQ0wAbM9pusiV3gIepTQHq2HKRKwgS~ypmv~0oqO5dDxaq9nhOWIBBmLAA49oHWuxUtnRe3sEYATwCwjIFEomhgJpg7vKGCF~87GdCV4JETW4PN2VvD4hsu2AVRd8yA~vUJvsXfhWwcElDeY4fsgpprl1lC3XGjJCnqdwbzRGPV5LNx5XEKLptcZtLeLcOmFJCvmVLe0bewJ7Vta0TSQM7T9AuklompjPPw__&Key-Pair-Id=APKAIE5G5CRDK6RD3PGA
